---
title: "Covid91 vaccine study Final2023"
author: "Erika Mondragon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



## Males
$H_0$:The covid921 vaccine does not exhibit a significant difference in effectiveness between male populations.

$H_A$: The covid921 vaccine does exhibit a significant difference in effectiveness between male populations

```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex == "M")
```

### Methods

Male is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results
```{r}
barchartGC(~infected + treatment ,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

The bar graph shows that the male recipients of the placebo had a greater rate of contracting COVID-19. In contrast to the male recipients of the 921 vaccination, who had a decreased risk of contracting COVID-19. Among the men who had covid 19, the male recipients of the COVID-19 vaccination, the risk of contracting the disease was significantly lower in those individuals. In contrast to the guys who were given a placebo, who were more likely to get COVID-19. In terms of the people who were never infected (OK), they have around similar percentages for both the placebo and vaccine. 

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

This table shows that those who were infected by covid 19 and recived the vaccince have a 30.43% of getting covid after reciveing the vaccine. Those who got the placebo have a 69.57% of contracting covid 19. From those who were never infeected (OK), the ones that got the vaccine had a 49.52% of getting covid 19, the ones who recieved the placebo had a 50.48%.

### Inferentail results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

On the chi-squared test the p-value is  0.0000000000001098, which is less than 0.05 so we reject the null hypothesis. On the fisher test, the p-value is 0.00000000000004631, we reject the null hypothesis, which is, the covid921 vaccine does not exhibit a significant difference in effectiveness between male populations. The odds ratio shows that you are 2.2 times more likely to get of covid if you are male and dont get the vaccine

## Females

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectiveness between female populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between female population

```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex == "F")
```

### Methods

Female is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results

```{r}
barchartGC(~infected + treatment ,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

This bar graph shows that in Females, those who were inflected by covid 19 and got the placebo have a higher chance of getting covid than those who were infected and got the vaccine. Among the women who had covid 19, the recipients of the covid 921 vaccination, the risk of contracting the disease was significantly lower in those individuals. In contrast to the guys who were given a placebo, who were more likely to get COVID-19. In terms of those who were never infected (OK), they have around similar percentages for both the placebo and vaccine. 

### Numerical results

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```

According to this table, those who were infected with COVID-19 and received the vaccination have a 40% chance of contracting the disease again after receiving the shot. The chance of getting COVID-19 is 60% for those who received the placebo. Among individuals who never contracted the virus (OK), those who received the vaccination had a 49.88% chance of contracting COVID-19, whereas those who received a placebo had a 50.12% chance.

### Inferentail results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Since the p-value for the chi-squared test is 0.0002732, which is less than 0.05, we may reject the null hypothesis. With a p-value of 0.0002162 on the Fisher test, we reject the null hypothesis, therefore, the covid921 vaccine does exhibit a significant difference in effectiveness between female population. According to the odds ratio, males who do not receive the vaccination have a 1.5 times increased risk of contracting COVID-19.

## LGBTQ

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectivness between LGBTQ populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between LGBTQ population

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ == "gay")
```
### Methods

LGBTQ is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results
```{r}
barchartGC(~infected + treatment ,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

This bar graph demonstrates that the members of LGBTQ who were infected with COVID-19 and received a placebo had a lower risk of contracting the disease than those who were infected and received the vaccination. The likelihood of getting COVID-19 was considerably greater in women who received the vaccine. As opposed to those who received a placebo, who had a lower risk of contracting COVID-19. Regarding individuals who never became infected (OK), the percentages for the vaccination and placebo are almost the same. 


### Numerical results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

This table indicates that after receiving the vaccination, those who were previously infected with COVID-19 had a 84.62% probability of getting infected again. The placebo group had a 15.38% probability of contracting COVID-19. Among people who never got covid 19 (OK), those who received the vaccine had a 49.98% risk of getting infected, whereas those who received a placebo had a 50.02% likelihood.


### Inferentail results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

Since the p-value for the chi-squared test is 0.00000006668, which is less than 0.05, we may reject the null hypothesis. With a p-value of 0.0000000109 on the Fisher test, we reject the null hypothesis, therefore,The covid921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ populations. The odds ratio shows that those who do receive the vaccination have a 0.181 times increased risk of contracting COVID-19.

## Druggies

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectiveness between druggies populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between druggies population

```{r}
library(dplyr)
Druggies = filter(FinalData, FinalData$DrugUser == "yes")
```

### Methods

Druggies is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat


### Results

### Graphical results

```{r}
barchartGC(~infected + treatment ,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

According to this graph, druggies who had the vaccine 921were more likely to contract COVID-19, contrast to individuals who received a placebo, who were at a decreased risk of developing COVID-19. Vaccines do not work on drug addicts. Both the placebo and vaccination groups had a similar chance of contracting COVID-19 among individuals who never had it.


### Numerical results

```{r}
table4 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table4)
colPerc(table4)
```

According to this data, the placebo group had an 11.67% risk of developing COVID-19. Contrasted with the 88.33% likelihood of getting covid among individuals who received the covid921. The individuals who received the placebo were far less likely than those who received the vaccination to get COVID-19. Those who never got covid, they both had an almost similar chance of getting covid, with the placebo having 51.19$ and the vaccine-48.81%.


### Inferentail results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

On the chi-squared test, the P value is 0.00000000284 ,which is less than 0.05, so we would reject the null hypothesis. On the fisher test, the P value is 0.000000000232,less than 0.05, so we reject the null hypothesis, The covid921 vaccine does not exhibit a significant difference in effectiveness between druggies populations. The odds ratio shows that it is 0.125 times more likely risk of contracting COVID-19 of those who do receive the vaccination.

# Overall Results and Conclusions

The original question was, is the covid 921 vaccine effective in the population subgroups by gender( males and females), drug use, and sexual orientation? In all of the population subgroups, we rejected the null hypothesis.

In the male population, since we reject the null hypothesis because the p-value is less than 0.05, based on the evidence we conclude that the covid921 vaccine does exhibit a significant difference in effectiveness between male populations.

For the female population, we also reject the null hypothesis because the p-value is less than 0.05, we conclude that the covid921 vaccine does exhibit a significant difference in effectiveness between the female population.

In the LGBTQ population, since we reject the null hypothesis because the p-value is less than 0.05, we accept that the covid921 vaccine does exhibit a significant difference in effectiveness between the LGBTQ population.

In the druggies population, since we reject the null hypothesis because the p-value is less than 0.05, we conclude that the covid921 vaccine does exhibit a significant difference in effectiveness between the druggies population.
