---
title: "Covid91 vaccine study Final2023"
author: "Erika Mondragon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



## Males
$H_0$:The covid921 vaccine does not exhibit a significant difference in effectiveness between male populations.
$H_A$: The covid921 vaccine does exhibit a significant difference in effectiveness between male populations

```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex == "M")
```

### Methods

Male is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results
```{r}
barchartGC(~infected + treatment ,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

### Inferentail results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
## Females

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectivness between female populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between female population

```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex == "F")
```

### Methods

Female is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results

```{r}
barchartGC(~infected + treatment ,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

### Numerical results

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```

### Inferentail results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

## LGBTQ

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectivness between LGBTQ populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between LGBTQ population

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ == "gay")
```
### Methods

LGBTQ is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat

### Results

### Graphical results
```{r}
barchartGC(~infected + treatment ,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

### Numerical results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

### Inferentail results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

## Druggies

$H_0$:The covid921 vaccine does not exhibit a significant difference in effectivness between druggies populations.

$H_A$:The covid921 vaccine does exhibit a significant difference in effectiveness between druggies population

```{r}
library(dplyr)
Druggies = filter(FinalData, FinalData$DrugUser == "yes")
```

### Methods

Druggies is the categorical variable, covid921 vaccine administartion is an explanatory variable. Covid921 effectiveness is a response variable is categorical. Therefore we use the helper file, Cat~Cat


### Results

### Graphical results

### Numerical results

### Inferentail results

# Overall Results and Conclusions